Mapping factors associated with adolescents' emotion dysregulation: A text-mining based systematic review

### 2020-06-07 manuscript.Rmd
Final screening done

keywords          : "keywords"
wordcount         : "X"
bibliography      : ["veni_sysrev.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

@@

<!-- From Barriola:
* adolescence, more than adulthood or childhood, is characterized by frequent and intense emotions (Larson et al. 1980; Larson and Lampmanpetraitis 1989).
* signiﬁcant psychosocial changes that call for an increase in sophistication of ER strategies (Steinberg 2005)
* novel stressors associated with social relationships, including pressures to conform to adult expectations relating to maturity, educational achievement, gender appropriate conduct, and other socially competent behaviors (Yap et al. 2007)
* Additionally, adolescents have an increased orientation toward autonomy (Spear 2000) including autonomous emotional regulation-->

@@

<!--
Barriola: What is most
* evident from the reviewed studies is that the samples
investigated have been mainly non-clinical, within the
early childhood period of development and have examined
predominantly maternal socialization factors. 
* the vast
majority have been restricted to toddlers and young chil-
dren and have only examined the inﬂuence of maternal,
and not paternal, expression.
Most of the literature has addressed emotion regulation in adult or infant samples. A smaller, but substantial, portion of the literature has examined emotion regulation in adolescents. 
-->

@@

The present paper aims to develop a nomological network of the constructs related to adolescent emotion dysregulation, based on a systematic review of the literature. A nomological network is a diagrammatic representation of a phenomenon; a proto-theoretical device that describes relationships between constructs relevant to the theory (REF: Cronbach & Meehl 1955; McKenna, 2006; Alavi 2018). It is a precursor to a more explicit theory of the phenomenon. The present paper uses an inductive approach to develop this nomological network. First, we conduct a systematic search to elicit a corpus of relevant literature. Second, we use text mining techniques to extract relevant constructs from the corpus. Third, we use a dictionary to pare down the extracted constructs to meaningful superordinate categories. Fourth, we map interrelations between these constructs. Fifth, we classify contstructs as potential predictors or outcomes, and we categorize predictors by level of analysis<!-- (CJ: Sameroff, which level of biopsychosocial)-->.
<!-- t, using text mining to map potential predictors of adolescent emotion dysregulation.
identify relevant constructs  of adolescents’ emotion regulation development, and moderators of parents’ role therein. The resulting insights are consolidated into theory.
The present study intends to map the extant literature 
First, I will systematically review predictors currently considered relevant in the adolescent emotion regulation literature.  -->

@@

All searches were conducted on Web of Science. The search strategy was based on procedures described by Staaks [@staaksSystematicReviewSearch]. First, we manually constructed a reference set of 15 articles. Then we constructed a search string to retrieve these articles from Web of Science. The search was overly inclusive, returning 29 records. As these were all highly relevant, we updated the reference set to include all 29 results. Next, we tested our search string, which consisted of synonyms of emotion regulation and adolescence: <!-- ### Reference set Our reference set is documented in `reference_set.bib`. -->
```
TS=("emotio* regulation" OR "anger regulation" OR "sadness regulation" OR "emotion* competence" OR "emotion* adjustment" OR "emotio* dysregulation" OR "anger dysregulation" OR "sadness dysregulation" OR "emotio* problem*" OR "emotion* maladjustment") AND TS=(adolescent* OR teen* OR pubert* OR youth)
```
This string returned 6653 results, and matched 25 of the 29 records in the reference set. Three search terms could be added to match all 29 reference set items. The terms `"emotio* socialization" OR "emotio* processes"`, as synonyms of emotion regulation, added 191 new hits. However, many of these hits were not directly relevant to emotion regulation. The term `"development*`, as synonym of adolescence, added 3628 new hits, many of which were not focused on the age range of adolescence. We thus deemed these terms to be overly inclusive, and proceeded with the original search string above.
```{r readrecs}
filenam <- "recs_6653.RData"
if(!file.exists(filenam)){
  f <- list.files("./recs/18-3-2020/", full.names = TRUE)
  f <- readFiles(f)
  recs <- convert2df(f)
  saveRDS(recs, filenam)
} else {
  recs <- readRDS(filenam)
}
```

@@

```{r removeduplicates}
filenam <- "unique_recs.RData"
if(!file.exists(filenam)){
  recs$doi <- recs$DI
  recs$doi[is.na(recs$doi)] <- paste0("fake", 1:sum(is.na(recs$doi)))
  if(any(duplicated(recs$doi))){
    recs$doi[duplicated(recs$doi)|duplicated(recs$doi, fromLast = T)]
    dup_doi <- duplicated(recs$doi)
    recs <- recs[!dup_doi, ]
  }
  recs$title <- recs$TI
  # locate and extract unique references
  potential_dups <- find_duplicates(recs, match_variable = "title", match_function = "stringdist", threshold = 5, to_lower = TRUE, remove_punctuation = TRUE)
  unique_recs <- extract_unique_references(recs, matches = potential_dups)
  saveRDS(unique_recs, filenam)
  write.csv(data.frame(dup_doi = sum(dup_doi), dup_title = (nrow(recs)-nrow(unique_recs))-sum(dup_doi)), "dups.csv")
} else {
  unique_recs <- readRDS(filenam)
  dups <- read.csv("dups.csv")
}
unique_recs$id_num <- 1:nrow(unique_recs)
```
Duplicates were identified based on exact DOI matches (`r dups[["dup_doi"]]` duplicates), and title similarity (`r dups[["dup_title"]]` duplicates). Manual screening in Rayyan resulted in the removal of an additional 13 duplicates.

@@

```{r}
head(unique_recs$SN)
df_screen <- unique_recs[, c("id_num", "TI", "AB")]
names(df_screen) <- c("id_num", "title", "abstract")
write.csv(df_screen, file = "asreview.csv", row.names = FALSE, fileEncoding = "UTF-8")
df_screen <- unique_recs[, c("id_num", "TI", "AU", "SO", "SN", "VL", "IS", "BP", "PY", "PU", "DI", "AB")]
names(df_screen) <- c("key", "title", "authors", "journal", "issn", "volume", "issue", "pages", "year", "publisher", "url", "abstract")
write.csv(df_screen, file = "rayyan.csv", row.names = FALSE)

```
We manually screened 500 abstracts for suitability in Rayyan. Then, we used ASReview to rank the predicted relevance of all papers based on the manually included and excluded articles. We screened the lowest-ranked articles to find an inflection point below which most articles were irrelevant, and excluded those from analysis.

@@

For this first analysis, we relied on author-provided keywords. We extracted keywords by document, and used a dictionary approach to classify closely related terms.

>>>>

keywords          : "keywords"
wordcount         : "X"
bibliography      : ["veni_sysrev.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
knit              : worcs:::cite_essential
---

@@

<!-- From Barriola:
* adolescence, more than adulthood or childhood, is characterized by frequent
and intense emotions (Larson et al. 1980; Larson and Lampmanpetraitis 1989).
* signi?cant psychosocial changes that call for an increase in sophistication
of ER strategies (Steinberg 2005)
* novel stressors associated with social relationships, including pressures to
conform to adult expectations relating to maturity, educational achievement,
gender appropriate conduct, and other socially competent behaviors (Yap et al.
2007)
* Additionally, adolescents have an increased orientation toward autonomy
(Spear 2000) including autonomous emotional regulation-->

@@

<!--
Barriola: What is most
* evident from the reviewed studies is that the samples
investigated have been mainly non-clinical, within the
early childhood period of development and have examined
predominantly maternal socialization factors.
* the vast
majority have been restricted to toddlers and young chil-
dren and have only examined the in?uence of maternal,
and not paternal, expression.
Most of the literature has addressed emotion regulation in adult or infant
samples. A smaller, but substantial, portion of the literature has examined
emotion regulation in adolescents.
-->

@@

The present paper aims to develop a nomological network of the constructs
related to adolescent emotion dysregulation, based on a systematic review of
the literature. A nomological network is a diagrammatic representation of a
phenomenon; a proto-theoretical device that describes relationships between
constructs relevant to the theory [see @alaviAligningTheoryMethodology2018;
@@cronbachConstructValidityPsychological1955;
@@mastersonNomologicalNetworkResearch2009]. It is a precursor to a more
explicit theory of the phenomenon. The present paper uses an inductive approach
to develop this nomological network. First, we conduct a systematic search to
elicit a corpus of relevant literature. Second, we use text mining techniques
to extract relevant constructs from the corpus. Third, we use a dictionary to
pare down the extracted constructs to meaningful superordinate categories.
Fourth, we map interrelations between these constructs. Fifth, we classify
contstructs as potential predictors or outcomes, and we categorize predictors
by level of analysis<!-- (CJ: Sameroff, which level of biopsychosocial)-->.
<!-- t, using text mining to map potential predictors of adolescent emotion
dysregulation.
identify relevant constructs of adolescents’ emotion regulation development,
and moderators of parents’ role therein. The resulting insights are
consolidated into theory.
The present study intends to map the extant literature
First, I will systematically review predictors currently considered relevant in
the adolescent emotion regulation literature.  -->

@@

All searches were conducted on Web of Science. The search strategy was based on
procedures described by Staaks [@staaksSystematicReviewSearch]. First, we
manually constructed a reference set of 15 articles. Then we constructed a
search string to retrieve these articles from Web of Science. The search was
overly inclusive, returning 29 records. As these were all highly relevant, we
updated the reference set to include all 29 results. Next, we tested our search
string, which consisted of synonyms of emotion regulation and adolescence: <!--
### Reference set Our reference set is documented in `reference_set.bib`. -->
```
TS=("emotio* regulation" OR "anger regulation" OR "sadness regulation" OR
"emotion* competence" OR "emotion* adjustment" OR "emotio* dysregulation" OR
"anger dysregulation" OR "sadness dysregulation" OR "emotio* problem*" OR
"emotion* maladjustment") AND TS=(adolescent* OR teen* OR pubert* OR youth)
```
This string returned 6653 results, and matched 25 of the 29 records in the
reference set. Three search terms could be added to match all 29 reference set
items. The terms `"emotio* socialization" OR "emotio* processes"`, as synonyms
of emotion regulation, added 191 new hits. However, many of these hits were not
directly relevant to emotion regulation. The term `"development*"`, as synonym
of adolescence, added 3628 new hits, many of which were not focused on the age
range of adolescence. We thus deemed these terms to be overly inclusive, and
proceeded with the original search string above.
```{r readrecs}
filenam <- "recs_6653.RData"
if(!file.exists(filenam)){
  f <- list.files("./recs/18-3-2020/", full.names = TRUE)
  f <- readFiles(f)
  recs <- convert2df(f)
  saveRDS(recs, filenam)
} else {
  recs <- readRDS(filenam)
}
```

@@

```{r removeduplicates}
filenam <- "unique_recs.RData"
if(!file.exists(filenam)){
  recs$doi <- recs$DI
  recs$doi[is.na(recs$doi)] <- paste0("fake", 1:sum(is.na(recs$doi)))
  if(any(duplicated(recs$doi))){
    recs$doi[duplicated(recs$doi)|duplicated(recs$doi, fromLast = T)]
    dup_doi <- duplicated(recs$doi)
    recs <- recs[!dup_doi, ]
  }
  recs$title <- recs$TI
  # locate and extract unique references
  potential_dups <- find_duplicates(recs, match_variable = "title", match_function = "stringdist", threshold = 5, to_lower = TRUE, remove_punctuation = TRUE)
  unique_recs <- extract_unique_references(recs, matches = potential_dups)
  saveRDS(unique_recs, filenam)
  write.csv(data.frame(dup_doi = sum(dup_doi), dup_title = (nrow(recs)-nrow(unique_recs))-sum(dup_doi)), "dups.csv", row.names = FALSE)
} else {
  unique_recs <- readRDS(filenam)
  dups <- read.csv("dups.csv")
}
unique_recs$id_num <- 1:nrow(unique_recs)

if(FALSE){
  # df_screen <- unique_recs[, c("id_num", "TI", "AB")]
  # names(df_screen) <- c("id_num", "title", "abstract")
  # write.csv(df_screen, file = "asreview.csv", row.names = FALSE, fileEncoding = "UTF-8")
  df_screen <- unique_recs[, c("id_num", "TI", "AU", "SO", "SN", "VL", "IS", "BP", "PY", "PU", "DI", "AB")]
  names(df_screen) <- c("key", "title", "authors", "journal", "issn", "volume", "issue", "pages", "year", "publisher", "url", "abstract")
  write.csv(df_screen, file = "rayyan.csv", row.names = FALSE)
}
rayyan_res <- read.csv("rayyan_exports/articles.csv")
rayyan_dups <- nrow(unique_recs) - nrow(rayyan_res)

```
Duplicates were identified based on exact DOI matches (`r dups[["dup_doi"]]` 
duplicates), and title similarity (`r dups[["dup_title"]]` duplicates). 
Manual screening in Rayyan resulted in the removal of an additional `r rayyan_dups` duplicates.

@@

```{r, message=FALSE}
screen <- read.csv("asreview_result_sysrevemotprob.csv")
screen <- screen[screen$key %in% rayyan_res$key, ]
screen$drop_these <- FALSE
screen$drop_these[which(screen$included == 1 | screen$rayyan == "False")] <- TRUE

if(!file.exists("recs_final.csv")){
  recs_final <- unique_recs[screen$key[!screen$drop_these], ]
write.csv(recs_final, "recs_final.csv", row.names = FALSE)
} else {
  recs_final <- read.csv("recs_final.csv")
}
```
Papers were screened based on two main criteria: First, papers had to address emotion regulation, emotional problems, or similar. Second, the target population must overlap (at least in part) with adolescence, defined as the age range from 10-25. Thus, for example, an article reviewing the role of hormonal changes in emotional adjustment throughout the entire life course, including adolescence, would be deemed relevant. Screening was initially conducted in Rayyan QRCI [@Ouzzani2016]. Papers were sorted using Rayyan's ranking algorithm. To train the algorithm on exclusion criteria, screening focused on abstracts most likely to be excluded. After coding `r sum(screen$rayyan == "True")+sum(screen$rayyan == "False")` papers, screening continued in ASReview - a free, open source alternative to Rayyan QRCI [@van_de_schoot_rens_2020_3828293]. This program uses text mining and active learning to build a customizable machine learning model for article inclusion. We used a naive Bayes model, and focused on exclusion of irrelevant papers. In ASReview, an additional `r sum(table(screen$included, screen$rayyan)[, 1])` papers were screened. Screening again focused on articles most likely to be excluded, and continued until, among the most recently screened 100 papers, only 6 were excluded. The screening procedure is detailed in Figure \@ref(prismachart). In total, `r nrow(recs_final)` papers were retained for analysis.

```{r prismachart, fig.cap="Record screening flowchart"}
library(tidySEM)

lo <- matrix(c("start", "dedup", "rayyan", "asreview", "included",
               "", "dedup2", "excluded", "", ""), nrow = 5, ncol = 2)
nodes <- data.frame(
  name = c("start", "dedup", "dedup2", "rayyan", "asreview", "included", "excluded"), 
  label = c(
    paste0("Records identified through\nWeb of Science (n = ", nrow(recs), ")"), 
    paste0("Records after duplicates removed\n(n = ", nrow(unique_recs), ")"), 
    paste0("Duplicates removed\n(n = ", nrow(recs)-nrow(screen), ")"), 
    paste0("Articles screened for eligibility\nusing Rayyan (n = ", nrow(screen), ")"), 
    paste0("Articles screened for eligibility\nusing ASReview (n = ", nrow(screen), ")"), 
    paste0("Studies included in\nqualitative synthesis (n = ", nrow(recs_final), ")"),
    paste0("Excluded studies (n = ", nrow(screen)-nrow(recs_final), ")"))
)
edges <- data.frame(
  from = c("start", "dedup", "rayyan", "asreview",
           "dedup", "rayyan", "rayyan", "asreview"),
  to = c("dedup", "rayyan", "asreview", "included",
         "dedup2", "dedup2", "excluded", "excluded"),
  label = c("",   "", sum(screen$rayyan == "True"), "",
            paste0("n = ", sum(dups)), 
            paste0("n = ", rayyan_dups),
            paste0("n = ", sum(screen$rayyan == "False")),
            paste0("n = ", sum(screen$included == 1, na.rm = TRUE)-sum(screen$rayyan == "False"))
  )
)
p <- graph_sem(layout = lo, nodes = nodes, edges = edges, angle = 1, rect_height = 1.2, rect_width = 1.3)
ggsave("prismachart.png", p, device = "png")
ggsave("prismachart.svg", p, device = "svg")
knitr::include_graphics("prismachart.svg")
```

@@

For this first analysis, we relied on author-provided keywords.
We extracted keywords by document, and used a dictionary approach
to classify closely related terms.
The dictionary was created specifically for this study.
The lead author composed a first draft of the dictionary based on prior knowledge of the field.
This draft was subsequently updated in two ways:
First, while writing the introduction for this paper, key terms from the theoretical and empirical literature were added to the dictionary.
Second, by including commonly occurring uncategorized words during preliminary analyses of the data.
An exclusion list of methodological terms and other non-substantive words was generated using the same procedure.
